Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2018-11-04 mprop/mprop.tex
Derivatives

\documentclass{mprop}
\usepackage{graphicx}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{url}
\usepackage{bm}
\usepackage{bbm}
\usepackage{booktabs}
%\usepackage{fancyvrb}
%\usepackage[final]{pdfpages}

@@

\newcommand{\dm}{\frac{\partial}{\partial\mathbf{m}}}
\newcommand{\dS}{\frac{\partial}{\partial\mathbf{S}}}
\newcommand{\da}{\frac{\partial}{\partial\alpha}}
\newcommand{\db}{\frac{\partial}{\partial\beta}}

@@

%state how you propose to solve the software development problem. Show that your
%proposed approach is feasible, but identify any risks.

In order to properly investigate the difference between variational inference
and maximum likelihood estimation for the model, we keep other parts of the
model the same. Namely, we set the covariance function to a version of the
automatic relevance detection kernel
\cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\[ k_\Theta(\mathbf{x}_i, \mathbf{x}_j) = \zeta\exp\left(
    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
where $\zeta$ is the overall ``scale'' factor for how similar or distant the
states are,
\[ \bm\Lambda = \diag(\lambda_1, \dots, \lambda_d) = \diag(\bm\lambda) \]
is a diagonal matrix that determines how relevant each feature is (where $d$
denotes the number of features), $\mathbbm{1}$ is defined as
\[ \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases} \]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). Thus we can set $\Theta = \{ \zeta,
\bm\lambda \}$. Similarly, we keep the expression for the prior of $\Theta$:
\begin{equation} \label{eq:theta_prior}
  p(\Theta | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
    \sum_{i=1}^d \log(\lambda_i + 1) \right).
\end{equation}

@@

% TODO: uncorrelated gammas instead of a multivariate Gaussian? review several options
In order to have a reasonable way of calculating/approximating the ELBO (see
Section \ref{vi_algs}), we need to have an approximating distribution for
$\Theta$, but unfortunately all the papers in Table \ref{table:approximations}
either fix it or treat it as a variational parameter. Hence we make our first
assumption without justification from previous literature:
\begin{equation} \label{eq:theta}
  q(\Theta) = q(\zeta) \times q(\bm\lambda).
\end{equation}
As it is reasonable for $\bm\lambda$ to take any value in $\mathbb{R}^d$, we can
approximate it simply as
\begin{equation} \label{eq:lambda}
  q(\bm\lambda) = \mathcal{N}(\bm\lambda; \bm\mu, \bm\Sigma)
\end{equation}
for some variational parameters $\bm\mu \in \mathbb{R}^d$ and $\bm\Sigma \in
\mathbb{R}^{d \times d}$. In contrast, as the covariance function must produce
non-negative values, we want to restrict $\zeta$ to positive real numbers. At
the same time, the distribution for $\zeta$ should have a flexible mean (i.e.,
not be tied to zero, like in the exponential distribution) and tails that
converge to zero as the value of the random variable moves away from the mean.
We might want to support some right skew, but the distribution should be close
to symmetric with at least some parameter values. This limits our choice of
distributions quite significantly, and we decide to go with the gamma
distribution as it is fairly flexible and commonly used
\cite{hogg2018introduction}. We then define the probability density function of
the gamma distribution as
\begin{equation} \label{eq:zeta}
  q(\zeta) = \Gamma(\zeta; \alpha, \beta) =
  \frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha - 1}e^{-\beta\zeta},
\end{equation}
where $\alpha > 0$ and $\beta > 0$ are two parameters of the distribution, and
$\Gamma(\cdot)$ is the gamma function. This gives us our vector of variational
parameters $\bm\nu = (\mathbf{m}, \mathbf{S}, \bm\mu, \bm\Sigma, \alpha,
\beta)^\intercal$.

@@

  \mathbb{E}\left[\log\left(\frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha -
        1}e^{-\beta\zeta}\right)\right] \\
  &- \mathbb{E}[\log\mathcal{N}(\bm\lambda; \bm\mu, \bm\Sigma)] &&-

@@

For the last two terms we can use the fact that, e.g., $-\mathbb{E}[\log

@@

distribution \cite{DBLP:journals/tit/AhmedG89}. Note that we cannot use this
fact for $\mathbb{E}[\log \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)]$  because
$\Kuu$ depends on $\Theta$. We also plug in the definitions of

@@

  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\bm\Sigma| + \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] - \sum_{i=1}^d \log(\lambda_i + 1)\right] \\

@@

  &- \mathbb{E}[\alpha\log\beta - \log\Gamma(\alpha) + (\alpha - 1)\log\zeta - \beta\zeta].

@@

$\mathbb{E}[\zeta] = \alpha/\beta$ and
$\mathbb{E}[\log \zeta] = \psi(\alpha) - \log \beta$, where $\psi$ is the

@@

\cite{DBLP:books/lib/Bishop07}. Finally, we can simplify

@@

@@

  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\bm\Sigma| + \frac{1}{2}\log|\mathbf{S}| - \frac{1}{2} \mathbb{E}[\Tr[\Kuu^{-2}]] -
    \sum_{i=1}^d \mathbb{E}[\log(\lambda_i + 1)] \\
  &- \frac{1}{2}\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] - \frac{1}{2}\mathbb{E}[\log|\Kuu|] + \alpha - \log\beta + \log\Gamma(\alpha) + (1 - \alpha)\psi(\alpha) \\

@@

\subsection{Derivatives}
% TODO: integrals always mean "on all range"
% TODO: some kind of intro

@@

We begin by removing terms independent of $\mathbf{m}$:
\[
  \begin{split}
    \frac{\partial\mathcal{L}}{\partial\mathbf{m}} =
    &-\frac{1}{2}\dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}]
    + \dm[\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}] \\
    &- \sum_{i=1}^N \sum_{t=1}^T \dm\mathbb{E}[V_r(s_{i,t})] -
      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
      s')\dm\mathbb{E}[V_r(s')],
  \end{split}
\]
where
\begin{equation} \label{eq:1}
\begin{split} % TODO: dominated convergence theorem
    \dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] &=
    \dm\iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\mathcal{N}(\mathbf{u};
    \mathbf{m},
    \mathbf{S})q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda \\
    &= \iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{m}} q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda.
  \end{split}
\end{equation}
$\frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{m}}$ can be calculated as
follows: % TODO: by the cookbook
\[
  \begin{split}
    \frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial\mathbf{m}} &=
    \mathcal{N}(\mathbf{u}; \mathbf{m},
    \mathbf{S})\dm\left[-\frac{1}{2}(\mathbf{u} -
      \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})\right] \\
    &= \mathcal{N}(\mathbf{u}; \mathbf{m},
    \mathbf{S})\left(-\frac{1}{2}\right)(\mathbf{S}^{-1} +
    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})\dm[\mathbf{u} -
    \mathbf{m}] \\
    &= \mathcal{N}(\mathbf{u}; \mathbf{m},
    \mathbf{S})\frac{1}{2}(\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u}
    - \mathbf{m}).
  \end{split}
\]
Plugging it back to \eqref{eq:1} gives:
\[
  \begin{split}
    \dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] &=
    \iiint(\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u})\frac{1}{2}(\mathbf{S}^{-1} +
    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})\mathcal{N}(\mathbf{u};
    \mathbf{m},
    \mathbf{S})q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda \\
    &=
    \frac{1}{2}\mathbb{E}[(\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u})(\mathbf{S}^{-1}
    + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})]
  \end{split}
\]

@@

Similarly to the previous section,
\[
  \begin{split}
    \frac{\partial\mathcal{L}}{\partial\mathbf{S}} &=
    \frac{1}{2}\dS[\log|\mathbf{S}|]
    -\frac{1}{2}\dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] \\
    &- \dS \left[\sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})] -
      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
      s')\mathbb{E}[V_r(s')] \right],
  \end{split}
\]
where % TODO: cite cookbook
\[ \dS[\log|\mathbf{S}|] =
  \frac{1}{|\mathbf{S}|}\frac{\partial|\mathbf{S}|}{\partial \mathbf{S}} =
  \frac{1}{|\mathbf{S}|}|\mathbf{S}|\mathbf{S}^{-\intercal} =
  \mathbf{S}^{-\intercal} \]
and
\[
  \begin{split}
    \dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] &=
    \dS\iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\mathcal{N}(\mathbf{u};
    \mathbf{m},
    \mathbf{S})q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda \\
    &=
    \iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\frac{\partial \mathcal{N}(\mathbf{u};
      \mathbf{m}, \mathbf{S})}{\partial \mathbf{S}}q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda.
  \end{split}
\]
$\frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{S}}$ can be simplified to
\[
  \begin{split}
    \frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{S}} &=
    \dS\left[\frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\exp \left( -\frac{1}{2}
        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
        \mathbf{m})\right)\right] \\
    &= \dS\left[\frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\right]\exp \left( -\frac{1}{2}
        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
        \mathbf{m})\right) \\
      &+ \frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\dS\left[\exp\left( -\frac{1}{2}
        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
        \mathbf{m})\right)\right] \\
    &=
    \frac{1}{(2\pi)^{m/2}}\dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right]\exp
    \left( -\frac{1}{2} (\mathbf{u} -
      \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})\right) \\
    &- \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})\dS[(\mathbf{u}
    - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})], \\
  \end{split}
\]
where
\[ \dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right] =
  -\frac{1}{2}|\mathbf{S}|^{-3/2}\frac{\partial |\mathbf{S}|}{\partial \mathbf{S}} =
  -\frac{1}{2}|\mathbf{S}|^{-3/2}|\mathbf{S}|\mathbf{S}^{-\intercal} =
  -\frac{1}{2|\mathbf{S}|^{1/2}}\mathbf{S}^{-\intercal} \]
and
\[ \dS[(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
  \mathbf{m})] = -\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
  \mathbf{m})^\intercal\mathbf{S}^{-\intercal} \]
according to \textit{The Matrix Cookbook} \cite{petersen2008matrix}. Plugging
the two expression back in gives
\[ \frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial
    \mathbf{S}} = -\frac{1}{2}\mathbf{S}^{-\intercal}\mathcal{N}(\mathbf{u};
  \mathbf{m}, \mathbf{S}) + \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m},
  \mathbf{S})\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
  \mathbf{m})^\intercal\mathbf{S}^{-\intercal}. \]
Thus
\[ \dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] =
  \frac{1}{2}\mathbb{E}[(\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u})(\mathbf{S}^{-\intercal}(\mathbf{u}
  - \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
  \mathbf{S}^{-\intercal})]. \]

\subsubsection{$\partial/\partial\alpha$}

>>>>

\documentclass{mprop}
\usepackage{graphicx}
\usepackage{amssymb,amsmath,amsthm,mathtools}
\usepackage{url}
\usepackage{bm}
\usepackage{bbm}
\usepackage{booktabs}
%\usepackage{fancyvrb}
%\usepackage[final]{pdfpages}

@@

\newcommand{\dm}{\frac{\partial}{\partial\mathbf{m}}}
\newcommand{\dS}{\frac{\partial}{\partial\mathbf{S}}}
\newcommand{\da}{\frac{\partial}{\partial\alpha_j}}
\newcommand{\db}{\frac{\partial}{\partial\beta_j}}

@@

%state how you propose to solve the software development problem. Show that your
%proposed approach is feasible, but identify any risks.
% TODO: Theta should be equal to lambda

In order to properly investigate the difference between variational inference
and maximum likelihood estimation for the model, we keep other parts of the
model the same. Namely, we set the covariance function to a version of the
automatic relevance detection kernel
\cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\[ k_\Theta(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
where $\lambda_0$ is the overall ``scale'' factor for how similar or distant the
states are,
\[ \bm\Lambda = \diag(\lambda_1, \dots, \lambda_d) \]
is a diagonal matrix that determines how relevant each feature is (where $d$
denotes the number of features), $\mathbbm{1}$ is defined as
\[ \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases} \]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). Thus we can set $\bm\lambda = (\lambda_0,
\dots, \lambda_d)^\intercal$, and $\Theta = \{ \bm\lambda \}$. Similarly, we
keep the expression for the prior of $\Theta$:
\begin{equation} \label{eq:theta_prior}
  p(\Theta | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
    \sum_{i=1}^d \log(\lambda_i + 1) \right).
\end{equation}

@@

In order to have a reasonable way of calculating/approximating the ELBO (see
Section \ref{vi_algs}), we need to have an approximating distribution for
$\Theta$, but unfortunately all the papers in Table \ref{table:approximations}
either fix it or treat it as a variational parameter. Hence we make our first
assumption without justification from previous literature:
\begin{equation} \label{eq:theta}
  q(\Theta) = \prod_{i=0}^dq(\lambda_i).
\end{equation}
We want to restrict $\lambda_0$ to be positive so that $k_\Theta$ would produce
non-negative values and not become trivial. Similarly, we want that $\lambda_i
\ge 0$ for $i = 1, \dots, d$ so that $\bm\Lambda$ is a positive-definite matrix.
Considering the possible distributions for all $d + 1$ variables, we would like
the mean to be flexible (i.e., not tied to zero, like with the exponential
distribution), and the tails to converge to zero as the value of the random
variable moves away from the mean. We might want to support some right skew, but
the distribution should be close to symmetric with at least some parameter
values. This limits our choice of distributions quite significantly, and we
decide to go with the gamma distribution as it is fairly flexible and commonly
used \cite{hogg2018introduction}. We then define the probability density
functions as
\begin{equation} \label{eq:zeta}
  q(\lambda_i) = \Gamma(\lambda_i; \alpha_i, \beta_i) =
  \frac{\beta_i^{\alpha_i}}{\Gamma(\alpha_i)}\lambda_i^{\alpha_i - 1}e^{-\beta_i\lambda_i}, \quad i = 0, \dots, d,
\end{equation}
where $\alpha_i > 0$ and $\beta_i > 0$ are parameters of the distribution, and
$\Gamma(\cdot)$ is the gamma function. This gives us our
vector of variational parameters $\bm\nu = (\mathbf{m}, \mathbf{S}, \alpha_0,
\beta_0, \dots, \alpha_d, \beta_d)^\intercal$.

@@

  \sum_{i=0}^d\mathbb{E}\left[\log\left(\frac{\beta_i^{\alpha_i}}{\Gamma(\alpha_i)}\lambda_i^{\alpha_i
        - 1}e^{-\beta_i\lambda_i}\right)\right] \\
  &-

@@

We can simplify the last term by noting that $-\mathbb{E}[\log

@@

distribution \cite{DBLP:journals/tit/AhmedG89}. Also note that we cannot use
this fact for $\mathbb{E}[\log \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)]$
because $\Kuu$ depends on $\Theta$. We also plug in the definitions of

@@

  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] - \sum_{i=1}^d \log(\lambda_i + 1)\right] \\

@@

  &- \sum_{i=0}^d \mathbb{E}[\alpha_i\log\beta_i - \log\Gamma(\alpha_i) + (\alpha_i - 1)\log\lambda_i - \beta_i\lambda_i]

@@

$\mathbb{E}[\lambda_i] = \alpha_i/\beta_i$ and
$\mathbb{E}[\log \lambda_i] = \psi(\alpha_i) - \log\beta_i$, where $\psi$ is the

@@

\cite{DBLP:books/lib/Bishop07}. Moreover, we can simplify

@@

Finally, as $\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}$ is a function of
$\mathbf{u}$ and $\Theta$, we can take the expectation of $\mathbf{u}$, leaving
the expectation of $\Theta$:
\[ \mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] =
  \mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}] +
  \mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] =
  \Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] +
  \mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}. \]

@@

  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| - \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-2}]] - \sum_{i=1}^d \mathbb{E}[\log(\lambda_i + 1)] - \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] \\
  &- \frac{1}{2}\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m} - \frac{1}{2}\mathbb{E}[\log|\Kuu|] + \sum_{i=0}^d \alpha_i - \log\beta_i + \log\Gamma(\alpha_i) + (1 - \alpha_i)\psi(\alpha_i) \\

@@

\subsection{Derivatives}
% TODO: integrals always mean "on all range"
% TODO: some kind of intro
% TODO: prove each application of the dominated convergence theorem

@@

We begin by removing terms independent of $\mathbf{m}$:
\[
  \begin{split}
    \frac{\partial\mathcal{L}}{\partial\mathbf{m}} =
    &- \frac{1}{2}\dm[\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}]
    + \dm[\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}] \\
    &- \sum_{i=1}^N \sum_{t=1}^T \dm\mathbb{E}[V_r(s_{i,t})] -
      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
      s')\dm\mathbb{E}[V_r(s')].
  \end{split}
\]
Here
\begin{align*}
  \dm[\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}] &= (\mathbb{E}[\Kuu^{-1}] + \mathbb{E}[\Kuu^{-1}]^\intercal)\mathbf{m}, \\
  \dm[\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}] &= \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}],
\end{align*}
and
\begin{equation} \label{eq:1}
  \begin{split}
    \dm\mathbb{E}[V_r(s)] &= \dm\iiint V_r(s) p(\mathbf{r} | \Theta,
    \mathbf{X_u}, \mathbf{u}) \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
    q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda \\
    &= \iiint V_r(s) p(\mathbf{r} | \Theta,
    \mathbf{X_u}, \mathbf{u}) \dm\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
    q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda,
  \end{split}
\end{equation}
where
\[
  \begin{split}
    \dm\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) &=
    \mathcal{N}(\mathbf{u}; \mathbf{m},
    \mathbf{S})\dm\left[-\frac{1}{2}(\mathbf{u} -
      \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})\right] \\
    &= \mathcal{N}(\mathbf{u}; \mathbf{m},
    \mathbf{S})\left(-\frac{1}{2}\right)(\mathbf{S}^{-1} +
    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})\dm[\mathbf{u} -
    \mathbf{m}] \\
    &= \mathcal{N}(\mathbf{u}; \mathbf{m},
    \mathbf{S})\frac{1}{2}(\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u}
    - \mathbf{m}).
  \end{split}
\]
Substituting it back into \eqref{eq:1} gives
\[
  \begin{split}
    \dm\mathbb{E}[V_r(s)] &= \frac{1}{2}\iiint V_r(s)
    (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})
    p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) \mathcal{N}(\mathbf{u};
    \mathbf{m}, \mathbf{S})
    q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda \\
    &= \frac{1}{2}\mathbb{E}[V_r(s) (\mathbf{S}^{-1} +
    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})].
\end{split}
\]
Hence
\[
  \begin{split}
    \frac{\partial\mathcal{L}}{\partial\mathbf{m}} =
    &-\frac{1}{2}(\mathbb{E}[\Kuu^{-1}] +
    \mathbb{E}[\Kuu^{-1}]^\intercal)\mathbf{m} +
    \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}] \\
    &- \frac{1}{2}\sum_{i=1}^N\sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})
    (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})] \\
    &- \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')
    \mathbb{E}[V_r(s') (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} -
    \mathbf{m})].
\end{split}
\]

@@

Similarly to the previous section,
\[
  \begin{split}
    \frac{\partial\mathcal{L}}{\partial\mathbf{S}} &=
    \frac{1}{2}\dS\log|\mathbf{S}|
    -\frac{1}{2}\dS\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] \\
    &- \sum_{i=1}^N \sum_{t=1}^T \dS\mathbb{E}[V_r(s_{i,t})] -
      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
      s')\dS\mathbb{E}[V_r(s')],
  \end{split}
\]
where
\[ \dS\log|\mathbf{S}| = \mathbf{S}^{-\intercal}, \]
and
\[ \dS\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] = \mathbb{E}[\Kuu^{-1}]^\intercal \]
by \emph{The Matrix Cookbook} \cite{petersen2008matrix}. Then
\[ \dS\mathbb{E}[V_r(s)] = \iiint V_r(s) q(\mathbf{r}) \dS\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
  q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda, \]
where
\[
  \begin{split}
    \dS\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) &=
    \dS\left[\frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\exp \left( -\frac{1}{2}
        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
        \mathbf{m})\right)\right] \\
    &= \dS\left[\frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\right]\exp \left( -\frac{1}{2}
        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
        \mathbf{m})\right) \\
      &+ \frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\dS\left[\exp\left( -\frac{1}{2}
        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
        \mathbf{m})\right)\right] \\
    &=
    \frac{1}{(2\pi)^{m/2}}\dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right]\exp
    \left( -\frac{1}{2} (\mathbf{u} -
      \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})\right) \\
    &- \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})\dS[(\mathbf{u}
    - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})]. \\
  \end{split}
\]
The two remaining derivatives can be taken with the help of \emph{The Matrix
  Cookbook} \cite{petersen2008matrix}:
\begin{align*}
  \dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right] =
  -\frac{1}{2}|\mathbf{S}|^{-3/2}\frac{\partial |\mathbf{S}|}{\partial \mathbf{S}} =
  -\frac{1}{2}|\mathbf{S}|^{-3/2}|\mathbf{S}|\mathbf{S}^{-\intercal} &=
  -\frac{1}{2|\mathbf{S}|^{1/2}}\mathbf{S}^{-\intercal}, \\
  \dS[(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
  \mathbf{m})] &= -\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
  \mathbf{m})^\intercal\mathbf{S}^{-\intercal}.
\end{align*}
Plugging them back in gives
\[ \dS\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) =
  -\frac{1}{2}\mathbf{S}^{-\intercal}\mathcal{N}(\mathbf{u};
  \mathbf{m}, \mathbf{S}) + \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m},
  \mathbf{S})\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
  \mathbf{m})^\intercal\mathbf{S}^{-\intercal}, \]
and
\[
  \begin{split}
    \dS\mathbb{E}[V_r(s)] &= \frac{1}{2}\iiint V_r(s)
    (\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
    \mathbf{m})^\intercal\mathbf{S}^{-\intercal} - \mathbf{S}^{-\intercal})
    q(\mathbf{r}) \mathcal{N}(\mathbf{u};
    \mathbf{m}, \mathbf{S}) q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda
    \\
    &= \frac{1}{2}\mathbb{E}[V_r(s)(\mathbf{S}^{-\intercal}(\mathbf{u} -
    \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
    \mathbf{S}^{-\intercal})].
  \end{split}
\]
Therefore
\[
  \begin{split}
    \frac{\partial\mathcal{L}}{\partial\mathbf{S}} &=
    \frac{1}{2}\mathbf{S}^{-\intercal} -
    \frac{1}{2}\mathbb{E}[\Kuu^{-1}]^\intercal - \frac{1}{2}
    \sum_{i=1}^N\sum_{t=1}^T
    \mathbb{E}[V_r(s_{i,t})(\mathbf{S}^{-\intercal}(\mathbf{u} -
    \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
    \mathbf{S}^{-\intercal})] \\
    &- \gamma\sum_{s' \in \mathcal{S}}\mathcal{T}(s_{i,t}, a_{i,t}, s')
    \mathbb{E}[V_r(s')(\mathbf{S}^{-\intercal}(\mathbf{u} -
    \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
    \mathbf{S}^{-\intercal})].
  \end{split}
\]

\subsubsection{$\partial/\partial\alpha_j$}

We begin in the usual way:
    \frac{\partial\mathcal{L}}{\partial\alpha_j} = &-
    \frac{1}{2}\da\mathbb{E}[\Tr[\Kuu^{-2}]] - \frac{1}{2}\da\mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}]] - \frac{1}{2}\da\mathbb{E}[\mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] - \frac{1}{2}\da\mathbb{E}[\log|\Kuu|] \\
    &+ \da\mathbb{E}[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}] +
    \da [\alpha_j + \log\Gamma(\alpha_j) + (1 - \alpha_j)\psi(\alpha_j)] \\
      &- \sum_{i=1}^N \sum_{t=1}^T \da\mathbb{E}[V_r(s_{i,t})] -
        \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
        s')\da\mathbb{E}[V_r(s')].
First,
